A survey on NoSQL and NewSQL datastores

### 2014-11-27 paper_dec/paper_dec.tex
introduction gefixed

%\documentclass{IEEEtran}
\documentclass{article}
\usepackage{cite}
\usepackage{url}
%\usepackage{lscape}

@@

Scientific progress has caused the cost of genome sequencing to drop at an exponential rate over the past decade and a half, even outpacing Moore's law since 2008 \cite{wetterstrand_sequencing_cost}. As in all sorts of biological, medical and pharmaceutical research, more and more genomes are being sequenced, the amount of data generated increases rapidly. The intermediary data generated in the sequencing pipeline for a human genome, ranges in the order of ?? TB. Also, a single human genome takes ?? MB/GB to store when sequenced,%TODO source
 so as this scales to genomes of millions of people and other organisms, storage requirements become ever more demanding in terms of scalability, latency and concurrency.
\\Increasing popularity of web services such as social networks has caused a similar explosion of data for web companies. To handle this so-called Big Data\cite{mashey1997big} in an adequate way, traditional relational DBMS no longer suffice. Therefor, large web companies such as Google and Amazon have developped new storage solutions that meet the demands for high scalability, low latency and high availability\cite{baker2011megastore}. This has spawned many so called NoSQL ('Not only SQL') databases, which loosen the rigid relational datamodel in favour of better scaling and easier distribution of the data.
\\NoSQL datastores come in multiple flavours and can be divided in a few categories based on the datamodel they use:
\begin{itemize}
\item Columnar stores: based on the datamodel pioneered by Google's BigTable, these store data in "a sparse, distributed, persistent multidimensional sorted map" \cite{chang2008bigtable}. In BigTable's case this is a map of row key, column key and a timestamp. In this way, multiple versions of the data can be stored in chronological order. Because the system doesn't interpret the data, relationships can't be modeled. This is left to application logic\cite{hecht2011nosql}.
\item Document stores: these encapsulate key-value pairs in JSON or JSON-like documents.
\end{itemize}

>>>>

\documentclass{IEEEtran}
%\documentclass{article}
\usepackage{cite}
\usepackage{url}
%\usepackage{lscape}

@@

Scientific progress has caused the cost of genome sequencing to drop at an exponential rate over the past decade and a half, even outpacing Moore's law since 2008 \cite{wetterstrand_sequencing_cost}. As in all sorts of biological, medical and pharmaceutical research, more and more genomes are being sequenced, the amount of data generated increases rapidly. The intermediary data generated in the sequencing pipeline for a human genome, ranges in the order of ?? TB. Also, a single human genome takes ?? MB/GB to store when sequenced,%TODO source
 so as this scales to genomes of millions of people and other organisms, storage requirements become ever more demanding in terms of scalability, latency and concurrency.
\\Increasing popularity of web services such as social networks has caused a similar explosion of data for web companies. To handle this so-called Big Data\cite{mashey1997big} in an adequate way, traditional relational DBMS no longer suffice. Therefor, large web companies such as Google and Amazon have developped new storage solutions that meet the demands for high scalability, low latency and high availability\cite{baker2011megastore}. This has spawned many so called NoSQL ('Not only SQL') databases, which loosen the rigid relational datamodel in favour of better scaling and easier distribution of the data.
\\NoSQL datastores come in multiple flavours and can be divided in a few categories based on the datamodel they use:
\begin{itemize}
\item Columnar stores: based on the datamodel pioneered by Google's BigTable, these store data in "a sparse, distributed, persistent multidimensional sorted map" \cite{chang2008bigtable}. In BigTable's case this is a map of row key, column key and a timestamp. In this way, multiple versions of the data can be stored in chronological order. Because the system doesn't interpret the data, relationships can't be modeled. This is left to application logic\cite{hecht2011nosql}.
\item Document stores: these store data as key-value pairs in documents. Values can be of a wide variety of types, such as nested documents, lists or scalar values. Attribute names can be dynamically specified at runtime and need not adhere to a fixed schema\cite{cattell2011scalable}. This is well suited for modelling complex data structures. Many document stores use the JSON-fileformat (or some derived form). In contrast to columnar stores, the values in documents are not opaque to the system and can thus be queried and indexed\cite{hecht2011nosql}.\\
The term NewSQL data stores is being used to classify a set of solutions that aim to combine the horizontal scalability and fault tolerance of NoSQL stores with the relational data model. Though they all use the relational model and offer SQL, NewSQL stores vary greatly under the hood, depending on the architecture they are built on\cite{grolinger2013data}. This paper proceeds first by elaborating on the methodology used to select and compare the chosen databases and then by reviewing each of the six datastores in detail.
\end{itemize}
